---
title: "DATA 607 Project 1"
author: "Samuel I Kigamba"
date: "September 22, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


####Project 1
In this project, you’re given a text file with chess tournament results where the information has some structure. Your
job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database)
with the following information for all of the players:
Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents
For the first player, the information would be:
Gary Hua, ON, 6.0, 1794, 1605
1605 was calculated by using the pre-tournament opponents’ ratings of 1436, 1563, 1600, 1610, 1649, 1663, 1716, and
dividing by the total number of games played.
If you have questions about the meaning of the data or the results, please post them on the discussion forum. Data
science, like chess, is a game of back and forth…
The chess rating system (invented by a Minnesota statistician named Arpad Elo) has been used in many other contexts,
including assessing relative strength of employment candidates by human resource departments.
You may substitute another text file (or set of text files, or data scraped from web pages) of similar or greater
complexity, and create your own assignment and solution. You may work in a small team. All of your code should be in
an R markdown file (and published to rpubs.com); with your data accessible for the person running the script.  


#Step 1
I downloaded the chess .txt file from the class link provided and uploaded it into my DATA 607 Project 1 repository on github. I then uploaded the file into R from Github and performed the below data manipulations using regular expressions.
Follow this link to view the original file <https://raw.githubusercontent.com/igukusamuel/DATA-607-Project-1/master/tournamentinfo.txt>

```{r}
library(RCurl)
chess <- getURL("https://raw.githubusercontent.com/nabilahossain/Class-IS607/master/Project%201/tournamentinfo.txt")
#chess
```

#Step 2
I used regular expressions to pull the (a) player's name and (b) the state.

```{r}
library(stringr)
name <- unlist(str_extract_all(chess, "\\| [[:upper:]- ]{4,} \\|"))
name <- str_replace_all(name, pattern = "(\\| )|([[:space:]]{1,}\\|)", replacement = "")
head(name[25:30])
state <- unlist(str_extract_all(chess, "\\|[[:space:]]{1,}[[A-Z]]{2} \\|"))
state <- str_replace_all(state, pattern = "(\\|[[:space:]]{1,})|([[:space:]]{1,}\\|)", replacement = "")
head(state, 10)
```

#Step 3
used regular expressions to extract (a) player's points and (b) pre-rating.

```{r}
total_points <- unlist(str_extract_all(chess, "\\|[[:digit:].[:digit:]]{3}[[:space:]]{1,}\\|"))
total_points <- str_replace_all(total_points, pattern = "(\\|)|([[:space:]]{1,}\\|)", replacement = "")
head(total_points, 10)
pre_rating <- unlist(str_extract_all(chess, "[:] [[:alnum:] ]{2,9}\\-\\>"))
pre_rating <- str_replace_all(pre_rating, pattern = "(\\: )|(\\s{1,}\\-\\>)|([O-Q]\\d{1,2})|(\\-\\>)", replacement = "")
pre_rating <- as.numeric(pre_rating)
head(pre_rating, 10)
```

#Step 4
We then create a table with (a) player's number, (b) name, (c) state, (d) total points and (e) pre-rating as the headers

```{r}
player_num <- unlist(str_extract_all(chess, "\\d{1,2}\\s\\|"))
player_num <- str_replace_all(player_num, pattern = "(\\s\\|)", replacement = "")
player_num <- as.numeric(player_num)
table1 <- data.frame(player_num = player_num, name = name, state = state, total_pts = total_points, pre_rating = pre_rating)
head(table1)
```

#Step 5
Extracting the first row from the txt file and creating a table named rounds

```{r}
r1 <- unlist(str_extract_all(chess, "\\d{1,2}\\s\\|(.+?)\\|[[:upper:]]{1}\\s{1,4}[[:digit:]]{0,3}\\|\n"))
r1 <- str_replace_all(r1, pattern = "(\\|[[:upper:]]{1} )", replacement = "\\|")
head(r1, 10)
r1 <- str_replace_all(r1, pattern = "([[:space:]]{2})|(\\|\n)", replacement = "")
rounds <- data.frame(str_split_fixed(r1, "\\|", 10))
names(rounds) <- c("num", "name", "total_pts", "round1", "round2", "round3", "round4", "round5", "round6", "round7")
tail(rounds, 4)
```

#Step 6
combine all the seven rounds of information extracted in table rounds in step 5 and create a second table named table2.

```{r}
library(reshape2)
r3 <- data.frame(rounds[c(1, 4:10)])
r3$num <- str_replace_all(r3$num, pattern = "\\s{1,}(\\d{1,2})", replacement = "\\1")
r4 <- melt(r3,id.vars="num", variable.name = "rounds", value.name = "opponent_number" )
tail(r4)
table2 <- subset(r4, opponent_number != " ")
table2$num <- as.numeric(table2$num)
tail(table2)
```

#Step 7
combine table 1 and table 2 to get the opponents pre-ratings and name it table 3

```{r}
library(sqldf)
table3 <- sqldf("select t2.num as 'player_num', t1.name as 'opponent_name', t2.rounds, t2.opponent_number, t1.pre_rating as 'opponent_pre_rating'   from table2 t2
        left join table1 t1
         on t2.opponent_number =  t1.player_num
        order by t2.num asc")
head(table3)
```

#Step 8
Aggregate the opponents pre-ratings into table 4

```{r}
library(stats)
table4 <- aggregate(opponent_pre_rating ~ player_num, data = table3, FUN = 'mean')
head(table4)
```

#Step 9
Join table 1 and table 4 to obtain a table of the required information and in the reqired format.


```{r}
Chess_Tournament <- sqldf("select t1.name as 'Player_Name', t1.state as 'Player_State', t1.total_pts as 'Total_Points', t1.pre_rating as 'Player_Pre-Rating', t4.opponent_pre_rating as 'Opponents_Average_Pre_Rating'
      from table1 t1 left join table4 t4
         on t4.player_num =  t1.player_num")
Chess_Tournament$Opponents_Average_Pre_Rating <- format(round(Chess_Tournament$Opponents_Average_Pre_Rating, 1), nsmall = 1)
head(Chess_Tournament, 30)
```

#Step 10
Finally save a .csv file into local drive or into github repository.

```{r}
write.csv(Chess_Tournament, file = "C:/Users/iguku/Google Drive/R and SQL/DATA 607 Project 1/DATA_607_Project_1.csv")
```